---
title: "Start"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: FALSE
    code_folding: show
    number_sections: TRUE
---

Same as weight-at-age, but simplified.

# SET-UP

```{r markdown, echo=F}
library(knitr)
opts_chunk$set(echo = T, collapse = T, fig.align = "center", fig.width = 9, fig.height = 6)
options(width = 300)
```    

```{r settings, message = F}
source('0.0_settings.R')

this.year <- 2022
```

# Input data
## Read
```{r read, message = F}
new <- FALSE
if(new){
    bio <- get.bio(species='maquereau',user=imlp.user,password=imlp.pass)
    save(bio,file=paste0('Rdata/bio_',Sys.Date(),'.Rdata'))
}else{
    df <- file.info(list.files("Rdata/", full.names = T,pattern="bio_"))
    load(rownames(df)[which.max(df$mtime)])
}

# Creation time
df <- file.info(list.files("Rdata/", full.names = T,pattern="bio_"))
df[which.max(df$mtime),'mtime']

```

## Subsetting and outlier removal
Could do outlier removal better.
```{r outliers, message = F}
# subset
bio.laa <- bio[!is.na(bio$length.frozen) & 
               !is.na(bio$agef) & 
                bio$agef>0 & 
                bio$agef<21,
               c('year','month','date','gear','nafo','weight','agef','sample.id','length.frozen')]
names(bio.laa)[7] <- 'age'
names(bio.laa)[9] <- 'length'

# remove outliers (see plot)
bio.laa$outlier <- FALSE
bio.laa$outlier <- ifelse(bio.laa$length<0.15,TRUE,bio.laa$outlier)              # less than 315cm
bio.laa$outlier <- ifelse(bio.laa$age>17,TRUE,bio.laa$outlier)                   # at least 4 out of 6 age 18 have wrong length AND weight
bio.laa[!bio.laa$outlier,] <- ddply(bio.laa[!bio.laa$outlier,],c('age'),transform,
                                    outlier=ifelse((length<=quantile(length,0.0001)|length>=quantile(length,0.9999)),TRUE,outlier))  # quick and dirty way to remove extremes

# cleaning
bio.laa$age.group <- ifelse(bio.laa$age < 10, bio.laa$age, 10)
```

## plots {.tabset}

### age vs length
```{r plot age_length}
ggplot(bio.laa,aes(x=age,y=length,col=outlier))+
    geom_point()
```

# Calculations
## Try different options 
Even trial run with different selectivity gears (despite not a great idea).
```{r options}
bio.laa <- bio.laa[!bio.laa$outlier,]
bio.laa$outlier <- NULL
bio.laa$length <- bio.laa$length*100 #cm

# 4TVWX + gillnets + june/july (IDEM laa!!!)
thisbio <- bio.laa[bio.laa$month %in% c(6,7) &
               bio.laa$nafo %in% c('4T', "4V", "4W", "4X") &
               bio.laa$gear %in% c("GND", "GNS", "GN"),]

laa <- ddply(thisbio,c('year','age.group'),summarise,
          l=mean(length),
          lsd=sd(length),
          n=length(length),
          N=length(unique(sample.id)))
laa$option <- '4TVWX_gill_JunJul'
names(laa)[2] <- 'age'

```


## plots {.tabset}

### laa
```{r plot comparison_laa}
ggplot(laa, aes(x=year, y=l, col=as.factor(age))) + 
    geom_line() + 
    labs(x='Year',y='Length (cm)',col='Age')+
    theme_minimal(base_size = 14) 
```

### number of fish
```{r plot comparison_nfish}
ggplot(laa, aes(x=year, y=age)) + 
    geom_tile(aes(fill=n)) + 
    geom_text(data=laa[laa$n<10,],aes(label=n))+
    labs(x='Year',y='Age',fill='n')+
    theme_minimal(base_size = 14)+
    scale_fill_viridis_c(direction = -1)+
    scale_x_continuous(expand=c(0,0))
```

### number of samples
```{r plot comparison_Nsample}
ggplot(laa, aes(x=year, y=age)) + 
    geom_tile(aes(fill=N)) + 
    geom_text(data=laa[laa$N<5,],aes(label=N))+
    labs(x='Year',y='Age',fill='N')+
    theme_minimal(base_size = 14)+
    scale_fill_viridis_c(direction = -1)+
    scale_x_continuous(expand=c(0,0))
```
# Fill and smooth 

Fit Noel Cadigan's state-space model for laa.
1) filter out noise.
2) fill in gaps.

Fits an AR1 process in 3 directions (age, year, cohort). Available in catchR package.

## fit model
```{r laa_fit, message = F, fig.height = 11,fig.width = 10,results = 'hide',message=FALSE}
laam <- laa[laa$n>=10,]
laam$cv <- with(laam,lsd/l)

all <- expand.grid(year=1968:this.year, age=min(laam$age):max(laam$age))
laam <- merge(all, laam, all.x = TRUE)

laa.fit <- armatrix.fit(year=laam$year,age=laam$age,x=laam$l,cv=laam$cv,shrink.cv = 0)
laa.fit  # convergence ok

save(laa.fit,file=paste0("Rdata/fitArmatrix_base_cv1shrink0_",Sys.Date(),".Rdata"))
```

## plots {.tabset}

### laa
```{r laa_final, message = F, fig.height = 3,fig.width = 5,results = 'hide',message=FALSE}
armatrix.waa(laa.fit)+scale_color_viridis_d()+scale_y_continuous(expand=c(0,0))
```

### Predicted obs
```{r laa_predobs, message = F, fig.height = 10,fig.width = 10,results = 'hide',message=FALSE}
armatrix.predobs(laa.fit,scale='free',ncol=3)
```

### Predicted
```{r laa_pred, message = F, fig.height = 10,fig.width = 10,results = 'hide',message=FALSE}
armatrix.pred(laa.fit,scale='free',ncol=3)
```

### Resid bubble
```{r laa_res, message = F, fig.height = 3,fig.width = 10,results = 'hide',message=FALSE}
armatrix.res(laa.fit)
```

### Resid dot
```{r laa_res2, message = F, fig.height = 10,fig.width = 10,results = 'hide',message=FALSE}
armatrix.res2(laa.fit)
```

### effects
```{r laa_effects, message = F, fig.height = 10,fig.width = 10,results = 'hide',message=FALSE}
armatrix.effects(laa.fit) # empty plot because of ggplotGrob function within (used to align plots)
```

### cvs
```{r laa_cv, message = F, fig.height = 10,fig.width = 10,results = 'hide',message=FALSE}
armatrix.cvs(laa.fit,scale='free',ncol=3)
```

# Save results
```{r laa_result, message = F}
laaf <- laa.fit$output
laaf <- laaf[,c(1:2,9)]
names(laaf)[3] <- 'length'

s <- dcast(laaf,year~age,value.var = 'length')
s[,2:ncol(s)] <- round(s[,2:ncol(s)] ,3)
write.csv(s, file=paste0('csv/laa_',this.year,'_base_cv1shrink0.csv'),row.names = FALSE)

p <- armatrix.waa(laa.fit,ylab='Length (cm)')+scale_color_viridis_d()
ggsave(filename = paste0('img/laa',this.year,'.png'),plot = p,units = 'cm',height = 8,width = 14)
```

# Plots
```{r laa_growthplot, message = F}
p2 <- ggplot(laaf,aes(x=age,y=length,col=year,group=year))+
    geom_line()+
    labs(y='Length (cm)',x='Age',col='Year')+
    scale_color_viridis_c()+
    scale_x_continuous(breaks=unique(laaf$age))
ggsave(filename = paste0('img/growth',this.year,'.png'),plot = p2,units = 'cm',height = 8,width = 14)
p2
```